---
title: "BioStat I Final Project"
author: "Jiefei Hu, Minhee Kwon, Yaoting Li, Xueqing Zhou"
date: "2022-12-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  error = FALSE,
  message = FALSE,
  warning = FALSE,
  tidy.opts = list(width.cutoff = 60)
)
```

```{r load_packages}
library(tidyverse)
library(gtsummary)
```

```{r load_data}
load("./brfss2013.RData")
dim(brfss2013)
head(brfss2013)
```

```{r data_clean}
# drop duplication
brfss_unique <- brfss2013[!duplicated(brfss2013$seqno), ]

# drop blank/refused/unknown/not sure
brfss_tmp <- brfss_unique %>%
  select(genhlth, sex, income2, weight2, height3, fruit1, fruitju1, avedrnk2, X_bmi5cat) %>%
  na.omit() %>%
  filter(!(genhlth == 7) & !(genhlth == 9)) %>%
  filter(!(income2 == 77) & !(income2 == 99)) %>%
  filter(!(weight2 == 7777) & !(weight2 == 9999) & !(weight2 == '')) %>%
  filter(!(height3 == 7777) & !(height3 == 9999)) %>%
  tibble()

# unify height unit in meters
tmp_height <- c()
for (i in brfss_tmp$height3) {
  if (as.numeric(i) <= 711) {
    tmp_height <- append(tmp_height, (as.numeric(substr(i, 1, 1))*30.48 + as.numeric(substr(i, 2, 3))*2.54)/100)
  } else {
    tmp_height <- append(tmp_height, as.numeric(substr(i, 2, 4))/100)
  }}
brfss_tmp$height_new <- tmp_height

# unify weight units in kilograms
tmp_weight <- c()
for (i in brfss_tmp$weight2) {
  if (as.numeric(i) <= 999) {
    tmp_weight <- append(tmp_weight, as.numeric(i)*0.4536)
  } else {
    tmp_weight <- append(tmp_weight, as.numeric(substr(i, 2, 4)))
  }}
brfss_tmp$weight_new <- tmp_weight

## categorized fruitju1 & fruit1
## merge 0 and 300 since sample size too small
categorized <- function(variable){
  ifelse(variable == 0 | variable == 300, "Rarely",
         ifelse(variable <= 199, "Always",
                ifelse(variable <= 299, "Very Often", "Sometimes")))
  }

# create new 'BMI' variable
# create new categorical fruit variable
brfss_tmp %>% 
  mutate(BMI = weight_new/height_new^2) %>%
  mutate(fruitju1_cat = as.factor(categorized(fruitju1)),
         fruit1_cat = as.factor(categorized(fruit1))) %>%
  select(genhlth, sex, income2, BMI, fruit1_cat, fruitju1_cat, avedrnk2, X_bmi5cat) -> brfss
```

## What is the outcome variable of interest in your analysis? 

Our outcome variable of interest is General Health 'genhlth'.

```{r outcome_variable_distribution}
plot(brfss$genhlth,
     main = "Distribution of General Health",
     xlab = "General Health")
```

##	What are the descriptive characteristics of your specified cohort?

### Distribution

```{r genhlth~sex}
barplot(table(brfss$sex, brfss$genhlth),
        main = "Distribution of General Health by Sex", xlab = "General Health",
        legend = rownames(table(brfss$sex, brfss$genhlth)), beside = TRUE)
```
```{r genhlth~income}
barplot(table(brfss$income2, brfss$genhlth),
        main = "Distribution of General Health by Income Level", xlab = "General Health",
        legend = rownames(table(brfss$income2, brfss$genhlth)), beside = TRUE)
```

```{r genhlth~BMI}
brfss %>%
  ggplot(aes(x = genhlth, y = BMI, fill = genhlth)) +
  geom_boxplot() + theme_classic() +
  labs(title = "Boxplot of General Health by BMI",
       x = "General Health", y = "BMI")
```

Table One is used to show numerical summaries of characteristics/variables of interest by the categorical variable General Health 'genhlth'. 

More specifically, it shows number of count and percentage for all categorical variables, and it presents median and IQR for all continuous variables.

```{r table_one}
data3 <- df %>%
  select(genhlth, avedrnk2, sex) %>%
  drop_na() %>%
  filter(avedrnk2<77)%>%
  group_by(genhlth) %>%
  summarize(avg_drink = mean(avedrnk2))


## generate overall table one
## test usage reference
## not sure if included BMI categories or not
## include p-value <= 0.2 (in this case all)
brfss %>%
  tbl_summary(by = genhlth,
              label = list(sex ~ "Sex",
                           income2 ~ "Income Level",
                           BMI ~ "BMI",
                           fruit1_cat ~ "Times Eating Fruit",
                           fruitju1_cat ~ "Times Drinking Fruit Juices",
                           avedrnk2 ~ "Number of Alcoholic Beverages Comsumption per Day",
                           X_bmi5cat ~ "BMI Category")) %>%
  add_p() %>%
  bold_labels()
```

```{r}
## table one categorized by sex
## Male
## error message since "underweight - fair" sample size too small: fix it?
brfss %>%
  filter(sex == "Male") %>%
  select(-sex) %>%
  tbl_summary(by = genhlth,
              label = list(income2 ~ "Income Level",
                           BMI ~ "BMI",
                           fruit1_cat ~ "Times Eating Fruit",
                           fruitju1_cat ~ "Times Drinking Fruit Juices",
                           avedrnk2 ~ "Number of Alcoholic Beverages Comsumption per Day",
                           X_bmi5cat ~ "BMI Category")) %>%
  add_p() %>%
  bold_labels()

## Female
brfss %>%
  filter(sex == "Female") %>%
  select(-sex) %>%
  tbl_summary(by = genhlth,
              label = list(income2 ~ "Income Level",
                           BMI ~ "BMI",
                           fruit1_cat ~ "Times Eating Fruit",
                           fruitju1_cat ~ "Times Drinking Fruit Juices",
                           avedrnk2 ~ "Number of Alcoholic Beverages Comsumption per Day",
                           X_bmi5cat ~ "BMI Category")) %>%
  add_p() %>%
  bold_labels()


## table one categorized by BMI category
## Obese
brfss %>%
  filter(X_bmi5cat == "Obese") %>%
  select(-X_bmi5cat) %>%
  tbl_summary(by = genhlth,
              label = list(sex ~ "Sex",
                           income2 ~ "Income Level",
                           BMI ~ "BMI",
                           fruit1_cat ~ "Times Eating Fruit",
                           fruitju1_cat ~ "Times Drinking Fruit Juices",
                           avedrnk2 ~ "Number of Alcoholic Beverages Comsumption per Day")) %>%
  add_p() %>%
  bold_labels()

##...other categories: do not need to compute every category, can choose some to analyze
```

```{r}
## overall table one
## more significant digits
descrTable(genhlth ~ ., data = brfss)

## filter male analysis
## similar for other categories
brfss_male <- brfss %>%
  filter(sex == "Male") %>%
  select(-sex)
descrTable(genhlth ~ ., data = brfss_male)
```

**TODO: fit regression model, generate graphs to visualize**

## sociodemographic variables

## Research Question 1
Is there an association between alcohol consumption and level of general health?
-  How does it apply to across income level, genders,etc?

```{r}
brfss_alcohol <- brfss_unique %>%
  select(genhlth, avedrnk2, sex) %>%
  drop_na() %>%
  filter(avedrnk2<77)%>%
  group_by(genhlth) %>%
  summarize(avg_drink = mean(avedrnk2))

ggplot(brfss_alcohol, aes(x= genhlth, y = avedrnk2)) + geom_point(aes(x= genhlth, y = avg_drink)) + labs(x = "general health", y = "average drink", title = "general health vs average drink") 

```
```{r}
## table of alcohol consumption, gender, general health
brfss_alcohol_sex <- brfss_unique %>%
  select(genhlth, avedrnk2, sex) %>%
  drop_na() %>%
  filter(avedrnk2<77)%>%
  group_by(genhlth, sex) %>%
  summarize(avg_drink = mean(avedrnk2))
brfss_alcohol_sex
```

```{r}
## table of alcohol consumption, gender, general thelah
ggplot(brfss_alcohol_sex, aes(x = genhlth, y = avg_drink, fill = sex)) +
         geom_bar(position = "dodge", stat = "identity") +
         labs(x = "General Health Status", y = "Daily Alcohol Consumption", fill = "Gender", title = "Relationship between genhealth and alcohol consumption") 
```
## Explanation: The plot shows that people in poor health drinks more alcohol than those people in good health. When comparing between genders, male consumes more alcohol than women

```{r}
## In progress
# t.test(avg_drink ~ genhlth, data = brfss_alcohol_sex)
```

