---
title: "BioStat I Final Project"
author: "Jiefei Hu, Minhee Kwon, Yaoting Li, Xueqing Zhou"
date: "2022-12-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  error = FALSE,
  message = FALSE,
  warning = FALSE,
  tidy.opts = list(width.cutoff = 60)
)
```

```{r load_packages}
library(tidyverse)
library(gtsummary)
```

```{r load_data}
load("./brfss2013.RData")
dim(brfss2013)
head(brfss2013)
```

## Research Question

What are the factors that will affect health conditions?

## What is the outcome variable of interest in your analysis? 

Our outcome variables of interest are variables measuring the health condition, which include General Health 'genhlth', Have any health care coverage 'hlthpln1', and How Much Time Do You Sleep 'sleptim1'.

### outcome variable distribution

```{r health_variables}
# drop duplication
brfss_unique <- brfss2013[!duplicated(brfss2013$seqno), ]

# select health condition measurement variables
brfss_hlth <- brfss_unique %>%
  select(genhlth, hlthpln1, sleptim1) %>%
  na.omit()
summary(brfss_hlth)
```

### outcome variable distribution graphs

```{r genhlth}
brfss_hlth %>%
  ggplot(aes(x = genhlth)) + 
  geom_bar(position = "stack") + 
  ylab("number of person") + 
  xlab("general health")
```

### Could someone do the graph for the distribution of other health variables?

1. hlthpln1

2. sleptim1

## What are the factors associated with health conditions?

```{r clean_factor_variables}
# drop refused/unknown/not sure
brfss_tmp <- brfss_unique %>%
  select(genhlth, hlthpln1, sleptim1, sex, income2, weight2, height3, fruit1, fruitju1, avedrnk2) %>%
  na.omit() %>%
  filter(sleptim1 <=24) %>%
  filter(!(income2 == 77) & !(income2 == 99)) %>%
  filter(!(weight2 == 7777) & !(weight2 == 9999) & !(weight2 == '')) %>%
  filter(!(height3 == 7777) & !(height3 == 9999)) %>%
  tibble()

# unify height unit in meters
tmp_height <- c()
for (i in brfss_tmp$height3) {
  if (as.numeric(i) <= 711) {
    tmp_height <- append(tmp_height, (as.numeric(substr(i, 1, 1))*30.48 + as.numeric(substr(i, 2, 3))*2.54)/100)
  } else {
    tmp_height <- append(tmp_height, as.numeric(substr(i, 2, 4))/100)
  }}
brfss_tmp$height_new <- tmp_height

# unify weight units in kilograms
tmp_weight <- c()
for (i in brfss_tmp$weight2) {
  if (as.numeric(i) <= 999) {
    tmp_weight <- append(tmp_weight, as.numeric(i)*0.4536)
  } else {
    tmp_weight <- append(tmp_weight, as.numeric(substr(i, 2, 4)))
  }}
brfss_tmp$weight_new <- tmp_weight

## categorized fruitju1 & fruit1
## merge 0 and 300 since sample size too small
categorized <- function(variable){
  ifelse(variable == 0 | variable == 300, "Rarely",
         ifelse(variable <= 199, "Always",
                ifelse(variable <= 299, "Very Often", "Sometimes")))
}

# create new 'BMI' variable
# create new categorical fruit variable
brfss_tmp %>% 
  mutate(BMI = weight_new/height_new^2) %>%
  mutate(fruitju1_cat = as.factor(categorized(fruitju1)),
         fruit1_cat = as.factor(categorized(fruit1))) %>%
  select(genhlth, hlthpln1, sleptim1, sex, income2, BMI, fruit1_cat, fruitju1_cat, avedrnk2) -> brfss
```

### Graphic summaries - relationships

```{r genhlth~sex}
barplot(table(brfss$sex, brfss$genhlth),
        main = "Distribution of General Health by Sex", xlab = "General Health",
        legend = rownames(table(brfss$sex, brfss$genhlth)), beside = TRUE)
```

```{r genhlth~income}
brfss %>%
  ggplot(aes(x = income2, fill = genhlth)) +
  geom_bar(position = "fill") +
  scale_fill_brewer(palette = "Greys") +
  ylab("Percentage of general health") +
  xlab("Income level")
```

```{r genhlth~BMI}
brfss %>%
  ggplot(aes(x = genhlth, y = BMI, fill = genhlth)) +
  geom_boxplot() + 
  labs(title = "Boxplot of General Health by BMI",
       x = "General Health", y = "BMI")
```

```{r genhlth ~ avedrnk}
# For alcohol, I'm not sure if log would help cuz it's too right-skewed
# Maybe just use a table to show the mean, which I put in the next chunk
brfss %>%
  ggplot(aes(x = genhlth, y = log(avedrnk2), fill = genhlth)) +
  geom_boxplot() + 
  labs(title = "Boxplot of General Health by average alcoholic drinking",
       x = "General Health", y = "Average alcoholic drinking")
```

```{r genhlth ~ avedrnk_mean}
brfss %>%
  group_by(genhlth) %>%
  summarize(mean_consumption = round(mean(avedrnk2), 2))
```

```{r genhlth ~ fruit1_cat}
# do we need to have some plot for fruit to show distribution?
```

```{r genhlth ~ fruitju1_cat}
# same question for fruitju
```

### Numeric summaries - Table One

Table One is used to show numerical summaries of characteristics/variables of interest by health variables.

It shows number of count and percentage for all categorical variables, and it presents mean and standard deviation for all continuous variables.

#### by General Health 'genhlth'

```{r table1_genhlth}
brfss %>%
  select(-hlthpln1, -sleptim1) %>%
  tbl_summary(by = genhlth,
              label = list(sex ~ "Sex",
                           income2 ~ "Income Level",
                           BMI ~ "BMI",
                           avedrnk2 ~ "Average alcoholic drinks per day",
                           fruit1_cat ~ "Frequency Eating Fruit",
                           fruitju1_cat ~ "Frequency Drinking Fruit Juices"),
              statistic = list(all_continuous() ~ "{mean}, ({sd})",
                               all_categorical() ~ "{n} ({p}%)"),
              digits = list(BMI ~ c(1,2),
                            avedrnk2 ~ c(1,2))) %>%
  add_p() %>%
  modify_header(label = "**Characteristics**") %>%
  bold_labels()
```

### by hlthpln1

```{r table1_hlthpln1}
brfss %>%
  select(-genhlth, -sleptim1) %>%
  tbl_summary(by = hlthpln1,
              label = list(sex ~ "Sex",
                           income2 ~ "Income Level",
                           BMI ~ "BMI",
                           avedrnk2 ~ "Average alcoholic drinks per day",
                           fruit1_cat ~ "Frequency Eating Fruit",
                           fruitju1_cat ~ "Frequency Drinking Fruit Juices"),
              statistic = list(all_continuous() ~ "{mean}, ({sd})",
                               all_categorical() ~ "{n} ({p}%)"),
              digits = list(BMI ~ c(1,2),
                            avedrnk2 ~ c(1,2))) %>%
  add_p() %>%
  modify_header(label = "**Characteristics**") %>%
  bold_labels()
```

### by sleep hours

```{r}
brfss %>%
  mutate(adequate_sleptim = if_else(sleptim1 < 7, "inadequate sleep", "adequate sleep")) -> brfss_slep
brfss_slep %>%
  select(-genhlth, -hlthpln1) %>%
  tbl_summary(by = adequate_sleptim,
              label = list(sex ~ "Sex",
                           income2 ~ "Income Level",
                           BMI ~ "BMI",
                           avedrnk2 ~ "Average alcoholic drinks per day",
                           fruit1_cat ~ "Frequency Eating Fruit",
                           fruitju1_cat ~ "Frequency Drinking Fruit Juices"),
              statistic = list(all_continuous() ~ "{mean}, ({sd})",
                               all_categorical() ~ "{n} ({p}%)"),
              digits = list(BMI ~ c(1,2),
                            avedrnk2 ~ c(1,2))) %>%
  add_p() %>%
  modify_header(label = "**Characteristics**") %>%
  bold_labels()
```

